---
title: "Lab 3"
author: "Aaron Hanrahan, Ritvik Maripally, Nick Cory"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(classdata)
library(ggplot2)
dat <- readxl::read_xls('C:/Users/Ritz/Desktop/DS202/GSS.xls')
```
##Renaming and cleaning
```{r}
dat<- rename(dat, 'year' = 'Gss year for this respondent')
dat<- rename(dat, 'happylvl' = 'General happiness')
dat<- rename(dat, 'polParty' = 'Political party affiliation')
dat<- rename(dat, 'income' = 'Respondents income')
dat<- rename(dat, 'gender' = 'Respondents sex')
dat<- rename(dat, 'education' = 'Highest year of school completed')
dat<- rename(dat, 'wedded' = 'Marital status')
dat<- rename(dat, 'hourperweek' = 'Number of hours usually work a week')
dat<- rename(dat, 'id' = 'Respondent id number')
dat<- rename(dat, 'ballot' = 'Ballot used for interview')

```

### Question 1

### Question 2

### Question 3
```{r}
happylvl <- filter(dat, happylvl == 'Very happy' | happylvl == 'Pretty happy')
p <- ggplot(happylvl, aes(x = happylvl, y = wedded)) + geom_count()
p + ggtitle('Marital Status vs. Happiness') + xlab('happylvl') + ylab('Marital Status')
# getting numerical count
happylvl %>% count(wedded)
```
It looks like(disregarding no answer) People who are married tend to be more pretty happy and very happy(aprox. 25k) while people separated,divorced, or widowed tend not to be happy as much.
### Question 4
```{r}
q = dat %>% group_by(`wedded`, `polParty`) %>%
  filter(`happylvl` == "Pretty happy" | `happylvl` == "Very happy") %>%
  summarize(`happylvl.n` = n()) %>%
  mutate(`happylvl.per2` = `happylvl.n`/ sum(`happylvl.n`))
q %>%
  ggplot(aes(x = `wedded`, y = `happylvl.per2`)) +
  geom_bar(stat = 'identity') + theme(axis.text.x = element_text(angle = 90)) +
  facet_wrap(~`polParty`)
```
Seems like Political party affiliation does play a part in happiness. Looks like people who are leaning towards democrats tend to have a higher happiness level while republicans and independents tend to have lower happiness levels than democrats.
### Question 5

### Question 6